---
title: "Russian River sockeye run timing 2019"
author: "Andy Barclay"
date: "February 19, 2020"
output: html_notebook
---
This R Markdown documents the genetic mixed stock analysis of sockeye salmon sampled from the lower Kenai River test fishery in 2018.  The purpose of this study is to determine if non-Russian River sockeye are present in the Kenai River during the early run.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Set up workspace
### Get functions and load tidy packages
```{r workspace setup}

#save.image("Russian_River_sockeye_run_timing_2019.RData") 
#load("Russian_River_sockeye_run_timing_2019.RData") 

source("C:\\Users\\awbarclay\\Documents\\R\\GitHubCloneFunctions.R")#GCL functions
source("V:\\Analysis\\Staff\\Andy Barclay\\R\\New Functions\\WorkspaceFolders.GCL.R")#A function I wrote

library("tidyverse")

```

### Create output folders 
```{r create folders, echo=TRUE, message=FALSE}

#WorkspaceFolders.GCL(Folders = c("data", "output", "bayes", "objects", "rubias"), Subfolders = list(bayes = c("baseline", "mixture", "control", "output"), rubias=c("baseline", "mixture", "output")), wd = getwd())

```

### Create locus control and get genotype data
```{r locus control, echo=TRUE}

#CreateLocusControl.GCL(markersuite = "Sockeye2011_96SNPs", username = "awbarclay", password = password)#Locus control

#LOKI2R.GCL(sillyvec="SKENNT19L", username = "awbarclay", password)#Pull Data from LOKI

```

### Create initial locus object
```{r initial locus object}

loci <- LocusControl$locusnames

```

## Data cleanup
### Check initial sample size
```{r initial sample size, echo=TRUE}

ColSize <- tibble(sillys = "SKENNT19L")
ColSize$Original_N <- SKENNT19L.gcl$n
ColSize

```

### Removing individuals with <80% of loci with gentotypes
Fish that have gentoypes for less than 80% of loci are likely to have poor quality DNA and might introduce gentotyping errors into the mixture data and reduce the accruacy of the mixed stock analysis (MSA).

3 fish were removed for missing genotypes
```{r missloci,echo=TRUE}

MissLOCI <- RemoveIndMissLoci.GCL(sillyvec = "SKENNT19L", proportion = 0.8)
MissLOCI

ColSize$AfterMissLoci_N <- SKENNT19L.gcl$n
ColSize

```

### Check for duplicate individuals and remove them
Fish with 99% of scores that match
No duplicates to remove
```{r dupckeck, echo=TRUE, message=FALSE}

dupcheck <- CheckDupWithinSilly.GCL(sillyvec = "SKENNT19L", loci = loci, quantile = NULL, minproportion = 0.99)$report
dupcheck

```
```{r final sample size}

ColSize$Final_N <- SKENNT19L.gcl$n
ColSize

```
## Combine MHCs and mtDNA loci
```{r combine loci}

CombineLoci.GCL(sillyvec = "SKENNT19L", markerset = loci[c(10, 12, 13)], update = TRUE)#mtDNA

CombineLoci.GCL(sillyvec = "SKENNT19L", markerset = loci[c(32, 33)], update = TRUE)#mtDNA

loci93 <- LocusControl$locusnames[-c(10, 12, 13, 32, 33)]

```


## Create BAYES baseline and get objects from baseline workspace
```{r get objects from baseline }

attach("V:/Analysis/2_Central/Sockeye/Cook Inlet/2012 Baseline/Baseline/EarlyRussianProoftests.RData")

groups <- groups

groupvec <- groupvec

Kenai14pops <- Kenai14pops

PopNames <- LocNames

basefortran <- CreateBaseline.GCL(sillyvec = Kenai14pops, loci = loci93, dir = "bayes/baseline", basename = "Kenai14pops96loci", type = "BAYES", groupvec = NULL)#BAYES

baseline <- create_rubias_baseline(sillyvec = Kenai14pops, loci = loci93, group_names = groups, groupvec = groupvec, baseline_name = "Kenai14pops96loci")#rubias

detach()

baseline
```
## Set up mixture objects
### Get metadata for SKENNT19L
```{r read in data sheet}

data_sheet <- read_csv(file="data/2019 Kenai Inriver Sockeye Genetics.csv") %>% 
  separate(Date, into = c("day", "month")) %>% 
  mutate(year = 2019, day = as.numeric(day), month = match(month, month.abb)) %>% 
  mutate(date = lubridate::ymd(paste(year, month, day, sep = "-"))) %>% 
  select(date, Lookup, `Select?`, Stratum, FK_FISH_ID)

data_sheet

```
### Get mixture strata from data sheet
```{r mixture strata}
require(lubridate)

strat_num <- data_sheet %>% 
  pull(Stratum) %>% 
  unique() %>% 
  na.omit() %>% 
  as.numeric()
  
  
mix_df <- data_sheet %>% 
  filter(`Select?` == "x") %>% 
  group_by(Stratum) %>% 
  summarize(start_date =  head(date, n =1), end_date = tail(date, n =1)) %>% 
  ungroup() %>% 
  mutate(mix_dates = paste0(month.abb[month(start_date)], day(start_date),"to", month.abb[month(end_date)], day(end_date)))

mixvec <- mix_df %>% 
  pull(mix_dates)

mix_df

```
### Get fish IDs for each mixture
```{r mixture IDs}
mix_ids <- lapply(mixvec, function(mix){
  
  dat <- filter(mix_df, mix_dates == mix);
  
  IDs <- data_sheet %>% 
  filter(`Select?` == "x", date%in%seq.Date(dat$start_date, dat$end_date, by = "day")) %>%
    pull(FK_FISH_ID) %>% 
    as.character()
  
  AttributesToIDs.GCL(silly = "SKENNT19L", attribute = "FK_FISH_ID", matching = IDs)
  
}) %>% 
  set_names(mixvec)

mix_ids

```

### Pool to mixture .gcl objects and check sample sizes
Mixture sample sizes range from 98 to 100 fish
```{r mixture setup}

require(lubridate)

for(mix in mixvec){

  ids <- mix_ids[[mix]]
  
  PoolCollections.GCL(collections = "SKENNT19L",
                      loci = loci93,
                      IDs = list(SKENNT19L = ids),
                      newname = mix
                      )
}

sapply(mixvec, function(mix){get(paste0(mix, ".gcl"))$n})#Mixture sample sizes

```
### Create mixture inputs
#### BAYES mixture files
```{r create BAYES mixture input files,echo=FALSE}

for(mix in mixvec){
  
  mixfortran <- CreateMixture.GCL(sillys = mix, loci = loci93, mixname = mix, dir = "bayes/mixture")
  
}

```
#### rubias mixture input files and object
```{r create rubias mixture input files,echo=FALSE}

mix.in <- create_rubias_mixture(sillyvec = mixvec, loci = loci93, path = "rubias/mixture")
mix.in

```

## Create BAYES Control Files
```{r bayes control files,echo=FALSE}

initmat <- MultiChainInits.GCL(npops = length(Kenai14pops), nchains=5, prop=.9)#Initial start values

priors <- Prior.GCL(groupvec = groupvec, groupweights = rep(1/max(groupvec), max(groupvec)), minval = 0.01)#Flat prior

for(mix in mixvec){
 
  CreateControlFile.GCL(sillyvec = Kenai14pops, loci = loci93, mixname = mix, basename = "Kenai14pops96loci", suffix = "", nreps = 40000, nchains = 5, groupvec = groupvec, priorvec = priors, initmat = initmat, dir = "bayes/control", seeds = matrix(sample(seq(10000),3*5), nrow = 3), thin=c(1,1,1), mixfortran = mixfortran, basefortran = basefortran, switches = "F T F T T T F")
  
}

```
### Move BAYES file to output folder for analysis
```{r move BAYES files}
#Move files into output folders for running BAYES

 folders <- c("control","mixture")
 dir = paste0(getwd(), "/bayes")
 
 bayesexe.paths = list.files("V:\\Analysis\\2_Central\\Chinook\\Cook Inlet\\2014\\Baseline\\BAYES\\BAYES exe files", full.names = T, include.dirs = T)
 
 bayesexe.files = list.files("V:\\Analysis\\2_Central\\Chinook\\Cook Inlet\\2014\\Baseline\\BAYES\\BAYES exe files", full.names = F, include.dirs = T)


#Move mixture and control files to output folder
 for(mix in mixvec){
   
   dir.create(path = paste0(dir, "/output/", mix))
 
    for(folder in folders){
          
          files = list.files(paste(dir, "/", folder, sep = ''), pattern = mix) 
          
          paths = paste(dir, "/", folder, "/", files, sep = '')
          
          for (path in paths){
            
            file.rename(from = path, to = sub(pattern = paste("bayes/", folder, sep = ''), x = path, replacement = paste("bayes/output/", mix, sep = ''), fixed = T))
                        
          }
    }
   
   file.copy(from = paste0(dir, "/baseline/Kenai14pops96loci.bse"), to = paste0(dir, "/output/", mix, "/Kenai14pops96loci.bse"))
   
   file.copy(from = bayesexe.paths, to = paste0("bayes/output/", mix, "/", bayesexe.files))
   
 }  


```
### Summarize BAYES mixture results
Not running analysis is BAYES right now. The files are available if needed.
```{r BAYES mixture summary by group, echo=FALSE, eval=FALSE}

# BAYES_results <- CustomCombineBAYESOutput.GCL(groupvec = seq(3), groupnames = groups, maindir = "BAYES/output", mixvec = Mixtures, prior = "", ext = "RGN", nchains = 5, burn = 0.5, alpha=0.1, PosteriorOutput = FALSE)

# results_sum_df<-BAYES_results%>% 
#   sapply(FUN = as_tibble,USE.NAMES = TRUE,simplify = FALSE) %>% 
#   bind_rows(.id="mixture") %>% 
#   mutate(group=rep(groups,length(Mixtures))) %>% 
#   select(mixture,group,mean,'5%','95%',sd)

# results_sum_df

```

### Checking compostion by population
```{r BAYES mixture summary by pop, echo=FALSE}

      
# BAYES_results_pop<-CustomCombineBAYESOutput.GCL(groupvec=1:14,groupnames=Kenai14pops, maindir="bayes/output", mixvec=Mixtures, prior="", ext="BOT", nchains=5, burn=0.5, alpha=0.1,PosteriorOutput=FALSE)
# 
# pop_results_sum_df<-BAYES_results_pop%>% 
#   sapply(FUN = as_tibble,USE.NAMES = TRUE,simplify = FALSE) %>% 
#   bind_rows(.id="mixture") %>% 
#   mutate(pop=rep(PopNames,length(Mixtures))) %>% 
#   select(mixture,pop,mean,'5%','95%',sd)
# 
# pop_results_sum_df

```

### Analyze mixtures with rubias
```{r rubias analysis, echo=FALSE}

priors_rubias <- tibble(collection = Kenai14pops, pi_param = priors, minval=0.01) 

run_rubias_mixture(reference=baseline, mixture=mix.in, group_names=groups, gen_start_col=5, method = "PB", 
                               alle_freq_prior = list(const_scaled = 1), pi_prior = priors_rubias , 
                               pi_init = NULL, reps = 25000, burn_in = 5000, pb_iter = 100, 
                               sample_int_Pi = 10, pi_prior_sum = 1, 
                               path = "rubias/output", seed = 56) 
 
```
### Summarize rubias results
```{r rubias mixture summary, echo = FALSE}

rubias_results_bias_cor <- custom_combine_rubias_output(rubias_output = NULL, mixvec = mixvec, group_names = groups, group_names_new = NULL, groupvec = NULL, groupvec_new = NULL, path = "rubias/output", alpha = 0.1, burn_in = 5000, bias_corr = TRUE, threshold = 5e-07, plot_trace = FALSE)

rubias_results_mcmc <- custom_combine_rubias_output(rubias_output = NULL, mixvec = mixvec, group_names = groups, group_names_new = NULL, groupvec = NULL, groupvec_new = NULL, path = "rubias/output", alpha = 0.1, burn_in = 5000, bias_corr = FALSE, threshold = 5e-07, plot_trace = FALSE)

results_all_methods <- bind_rows(
  
  rubias_results_bias_cor %>% 
  mutate(method = "rubias_PB", lo5CI = `5%`, hi95CI = `95%`) %>% 
  select(-`P=0`, -`5%`, -`95%`, -median),

rubias_results_mcmc %>% 
  mutate(method = "rubias_MCMC", lo5CI = `5%`, hi95CI = `95%`)%>% 
   select(-`P=0`, -`5%`, -`95%`, -median),

# results_sum_df %>% 
#   mutate(mixture_collection = mixture, repunit = factor(group,levels=groups), method = "BAYES", lo5CI = `5%`, hi95CI = `95%`)%>% 
#   select(-group, -mixture, -`5%`, -`95%`)

) %>% 
  mutate(mixture_collection = factor(mixture_collection, levels = mixvec))

results_all_methods
```
### Plot results
bias correction doesn't change the results much; go with the MCMC estimates
```{r plot estimates to compare, fig.height=10, fig.width=8}

GrCol <- c("blue","red","green")

results_all_methods %>% 
  ggplot(aes(y = mean, x = repunit, fill = repunit))+
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = lo5CI, ymax = hi95CI, width = 0.3), position = "dodge")+
  scale_fill_manual(name = "Reporting Group", values = GrCol)+
  facet_grid(mixture_collection~method) +
  ylab("Proportion (90% CI)")+
  xlab("Reporting Group")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.text.x = element_text(size = 14), 
          strip.text.y = element_text(size = 14), 
          axis.title.x = element_text(size = 13),
          axis.title.y = element_text(size = 13),
          legend.position = "none")
  
  
ggsave(filename="output/2018_RussianRiverRunTimingStockComps_AllMethods.pdf",device = "pdf")

```

## Analyze mixtures in rubias with finer-scale reporting groups
```{r fine scale reporting groups}

groupvec6 <- c(1, 1, 1, 1, 1, 2, 3, 4, 4, 4, 4, 2, 5, 6)
groups6 <- c("UpperTribs","UpperMainstem","RussianEarly","RussianLate","HiddenL","SkilakOutlet")

attach("V:/Analysis/2_Central/Sockeye/Cook Inlet/2012 Baseline/Baseline/EarlyRussianProoftests.RData")

baseline_6groups <- create_rubias_baseline(sillyvec = Kenai14pops, loci = loci93, group_names = groups6,groupvec = groupvec6, baseline_name = "Kenai14pops96loci6groups")#rubias

detach()

baseline_6groups

```
```{r rubias fine-scale analysis, echo=FALSE}

dir.create(path = "rubias/output/6_groups")

sapply(mixvec, function(mix){
  
  run_mix<-mix.in %>% filter(collection == mix)
  
  run_rubias_mixture(reference = baseline_6groups, mixture = run_mix, group_names = groups6, gen_start_col = 5, method = "MCMC", alle_freq_prior = list(const_scaled = 1), pi_prior =priors_rubias , pi_init = NULL, reps = 25000, burn_in = 5000, pb_iter = 100, sample_int_Pi = 10, pi_prior_sum = 1, path = "rubias/output/6_groups", seed = 56) 
  
}) 

```
### Summarize rubias results
```{r rubias mixture summary,echo=FALSE}


rubias_results_mcmc_6groups <- custom_combine_rubias_output(rubias_output = NULL, mixvec = mixvec, group_names = groups6, group_names_new = NULL, groupvec = NULL, groupvec_new = NULL, path = "rubias/output/6_groups", alpha = 0.1, burn_in = 5000, bias_corr = FALSE, threshold = 5e-07, plot_trace = FALSE)

results_6groups <- rubias_results_mcmc_6groups %>% 
  mutate(method = "rubias_MCMC",lo5CI = `5%`, hi95CI = `95%`)%>% 
   select(-`P=0`,-`5%`,-`95%`,-median) %>% 
  mutate(mixture_collection = factor(mixture_collection, levels = mixvec))

results_6groups
```
### Plot results for 6 groups
```{r plot estimates to compare, fig.height=10, fig.width=8}

GrCol6 <- c("magenta","cyan","red","green","orange","grey")

results_6groups %>% 
  ggplot(aes(y=mean,x=repunit,fill=repunit))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.text.x=element_text(size=14), 
          strip.text.y=element_text(size=14), 
          axis.title.x=element_text(size=13),
          axis.title.y=element_text(size=13),
          legend.position="none")+
  geom_bar(stat="identity",position = "dodge") +
  geom_errorbar(aes(ymin = lo5CI, ymax = hi95CI, width = 0.3), position = "dodge")+
  scale_fill_manual(name = "Reporting Group", values = GrCol6)+
  facet_grid(~mixture_collection) +
  ylab("Proportion (90% CI)")+
  xlab("Reporting Group")
  
ggsave(filename="output/2019_RussianRiverRunTimingStockComps_AllMethods_6groups.pdf",device = "pdf")

```
### Plot rubias MCMC results for 3 and 6 groups
```{r plot estimates to compare, fig.height=10, fig.width=8}

GrCol_all <- c("red", "green", "blue", "magenta", "cyan", "orange", "grey")

levels7 <- c("RussianEarly", "RussianLate", "KenaiOther", "UpperTribs", "UpperMainstem", "HiddenL", "SkilakOutlet")

df <- bind_rows( results_all_methods %>% 
                  filter(method == "rubias_MCMC") %>% 
                  mutate(n_groups="3 groups"),
                
                results_6groups %>% 
                  filter(method == "rubias_MCMC") %>% 
                  mutate(n_groups = "6 groups") 
) %>% 
  group_by(mixture_collection, n_groups) %>% 
  arrange(desc(n_groups), match(mixture_collection, mixvec), match(repunit, levels7)) %>% 
  ungroup() %>% 
  mutate(repunit = factor(repunit, levels = levels7))
  

  df %>% 
  ggplot(aes(y = mean, x = repunit, fill = repunit))+
    ggtitle(label = "2019 Russian River run timing stock composition estimates") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.text.x = element_text(size = 14), 
          strip.text.y = element_text(size = 14), 
          axis.title.x = element_text(size = 13),
          axis.title.y = element_text(size = 13),
          legend.position = "none")+
  geom_bar(stat = "identity",position = "dodge") +
  geom_errorbar(aes(ymin = lo5CI, ymax = hi95CI, width = 0.3), position = "dodge")+
  scale_fill_manual(name = "Reporting Group", values = GrCol_all)+
  facet_grid(mixture_collection~n_groups, scales = "free_x", space = "free")+
  ylab("Proportion (90% CI)")+
  xlab("Reporting Group")
  
ggsave(filename = "output/2019_RussianRiverRunTimingStockComps_rubias_MCMC_3vs6_groups.pdf", device = "pdf")


```
```{r plot rubias 3 group results, fig.height=3, fig.width=6}

df %>% 
  filter(n_groups == "3 groups") %>% 
  ggplot(aes(y = mean, x = repunit, fill = repunit))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.text.x = element_text(size = 14), 
          strip.text.y = element_text(size = 14), 
          axis.title.x = element_text(size = 13),
          axis.title.y = element_text(size = 13),
          legend.position = "none")+
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = lo5CI, ymax = hi95CI, width = 0.3), position = "dodge")+
  scale_fill_manual(name = "Reporting Group", values = GrCol_all)+
  facet_wrap(~mixture_collection, nrow = 1)+
  ylab("Proportion (90% CI)")+
  xlab("Reporting Group")

ggsave(filename = "output/RussianR_run_timing_2019_results_plot.pdf", device = "pdf")
 
```

### Table the estimates
```{r table estimates}

df %>% 
  filter(n_groups == "3 groups") %>% 
  select(mixture_collection, repunit  ,mean,lo5CI,hi95CI,sd) %>%
  write_excel_csv(path="output/Rubias_MCMC_3groups_estimates.csv")

df %>% 
  filter(n_groups=="6 groups") %>% 
  select(mixture_collection,repunit,mean,lo5CI,hi95CI,sd) %>%
  write_excel_csv(path="output/Rubias_MCMC_6groups_estimates.csv")

```

